### 2019-08-08 paper/results.tex
adding velocity dispersion as color

If we assume that the rise in velocity dispersions at cooler effective
temperatures is caused by an inaccurate period-color relation rather than
mass-dependent dynamical heating, we can estimate what the shape of the
period-color relations {\it should} be.
Figure \ref{fig:age_cut} indicates that the period-color relation flattens
out, so we applied the same analysis to groups of stars with similar rotation
periods, equivalent to a completely flat period-color relation.
The top panel of figure \ref{fig:period_cut} shows the \mct\ sample with stars
in different period ranges plotted in different colors.
The bottom panel shows the velocity dispersion of each group as a function of
effective temperature.
\begin{figure}
  \caption{
This figure is similar to figure \ref{fig:age_cut}, with stars divided into
    period groups rather than age groups.
The velocity dispersion is more constant across effective temperatures for the
    most slowly rotating stars, compared to the stars selected with the
    \citet{angus2019} gyrochronology model, indicating that the gyrochronology
    models flatten out, and possibly even invert, at old ages.
}
  \centering
    \includegraphics[width=1\textwidth]{period_cut}
\label{fig:period_cut}
\end{figure}
Once again, the velocity dispersion increases with rotation period overall.
For the most rapidly rotating groups of stars, velocity dispersion decreases
with \teff\ as expected given the positively sloped period-color relation of
Praesepe and other young clusters: late K dwarfs rotate more slowly than early
K dwarfs of the same age.
Between rotation periods of 15 and 25 days, the temperature dependence of the
velocity dispersion starts to disappear, indicating that the period-color
relation becomes flat: late K dwarfs rotate at the same rate as early K dwarfs
of the same age.
At long rotation periods, the velocity dispersion still increases with
effective temperature, although the increase is much more modest compared to
the oldest stars in figure \ref{fig:age_cut}.
This suggests that rotation period may {\it decrease} with effective
temperature, \ie\ late K dwarfs may rotate {\it more} rapidly than early K
dwarfs.
This would be a paradigm shift for gyrochronology, stellar spin-down rate is
thought to be directly tied to magnetic field strength.
The deeper convection zones of later type stars generate stronger magnetic
fields which {\it should} lead to more efficient angular momentum loss.
Unfortunately, we still cannot rule out the possibility that mass-dependent
heating is acting on these stars which could be responsible for some, if not
all, of the increased velocity dispersion at cooler effective temperatures.

>>>>

If we assume that the rise in velocity dispersions at cooler effective
temperatures is caused by an inaccurate period-color relation rather than
mass-dependent dynamical heating, we can estimate what the shape of the
period-color relations {\it should} be.
Figure \ref{fig:age_cut} indicates that the period-color relation flattens
out, so we applied the same analysis to groups of stars with similar rotation
periods, equivalent to a completely flat period-color relation.
The top panel of figure \ref{fig:period_cut} shows the \mct\ sample with stars
in different period ranges plotted in different colors.
The bottom panel shows the velocity dispersion of each group as a function of
effective temperature.
\begin{figure}
  \caption{
This figure is similar to figure \ref{fig:age_cut}, with stars divided into
    period groups rather than age groups.
The velocity dispersion is more constant across effective temperatures for the
    most slowly rotating stars, compared to the stars selected with the
    \citet{angus2019} gyrochronology model, indicating that the gyrochronology
    models flatten out, and possibly even invert, at old ages.
}
  \centering
    \includegraphics[width=1\textwidth]{period_cut}
\label{fig:period_cut}
\end{figure}
Once again, the velocity dispersion increases with rotation period overall.
For the most rapidly rotating groups of stars, velocity dispersion decreases
with \teff\ as expected given the positively sloped period-color relation of
Praesepe and other young clusters: late K dwarfs rotate more slowly than early
K dwarfs of the same age.
Between rotation periods of 15 and 25 days, the temperature dependence of the
velocity dispersion starts to disappear, indicating that the period-color
relation becomes flat: late K dwarfs rotate at the same rate as early K dwarfs
of the same age.
At long rotation periods, the velocity dispersion still increases with
effective temperature, although the increase is much more modest compared to
the oldest stars in figure \ref{fig:age_cut}.
This suggests that rotation period may {\it decrease} with effective
temperature, \ie\ late K dwarfs may rotate {\it more} rapidly than early K
dwarfs.
This would be a paradigm shift for gyrochronology, stellar spin-down rate is
thought to be directly tied to magnetic field strength.
The deeper convection zones of later type stars generate stronger magnetic
fields which {\it should} lead to more efficient angular momentum loss.

\begin{figure}
  \caption{
The rotation periods vs effective temperatures of stars in the \mct\ sample,
    colored by the velocity dispersions of stars in restricted period and
    temperature ranges.
}
  \centering
    \includegraphics[width=1\textwidth]{dispersion_period_teff}
\label{fig:dispersion_period_teff}
\end{figure}

Figure \ref{fig:dispersion_period_teff} shows the stars in our sample on a
plot of period vs effective temperature, coloured by their (\vb) velocity
dispersion, where velocity dispersion was calculated in period and temperature
bins (made apparent by the visible grid-like structure).
If we assume that mass dependent heating does {\it not} affect this sample,
velocity dispersion can be interpreted as an age proxy and stars bins of
similar colour in figure \ref{fig:dispersion_period_teff} are similar ages.
Interpreted this way, it appears as though isochrones slope downwards in
period-temperature space at old ages.
This figure also suggests that the oldest stars with measured rotation periods
are cool, which may be because cooler stars remain active for longer, and
their rotation periods can therefore be measured at older ages.

Unfortunately, we still cannot rule out the possibility that mass-dependent
heating is acting on these stars which could be responsible for some, if not
all, of the increased velocity dispersion at cooler effective temperatures.
